---
output: 
  rmarkdown::github_document:
    df_print: kable
---
```{r pkg-knitr-opts, include=FALSE}
hrbrpkghelpr::global_opts()
```

```{r badges, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::stinking_badges()
```

```{r description, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::yank_title_and_description()
```

## NOTE

I don't think this will work on Windows.

## What's Inside The Tin

The following functions are implemented:

```{r ingredients, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::describe_ingredients()
```

## Installation

```{r install-ex, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::install_block()
```

## Usage

```{r usage, cache=FALSE}
library(xattrs)
library(tidyverse) # for printing

# current verison
packageVersion("xattrs")

```

### Basic Operation

Extended attributes seem to get stripped when R builds pkgs so until I can figure out an easy way not to do that, just find any file on your system that has an `@` next to the permissions string in an `ls -l` directory listing.

```{r 1, eval=FALSE}
sample_file <- "~/Downloads/AdvancedTechniquesWithAWSGlueETLJobs.pdf"

list_xattrs(sample_file)
## [1] "com.apple.metadata:kMDItemWhereFroms"
## [2] "com.apple.quarantine"  

get_xattr_size(sample_file, "com.apple.metadata:kMDItemWhereFroms")
## 157
```

Extended attributes can be _anything_ so it makes alot of sense to work with the contents as a raw vector:

```{r 2, eval=FALSE}
get_xattr_raw(sample_file, "com.apple.metadata:kMDItemWhereFroms")
##   [1] 62 70 6c 69 73 74 30 30 a2 01 02 5f 10 6b 68 74 74 70 73 3a 2f
##  [22] 2f 66 69 6c 65 73 2e 73 6c 61 63 6b 2e 63 6f 6d 2f 66 69 6c 65
##  [43] 73 2d 70 72 69 2f 54 33 56 31 5a 44 51 48 4d 2d 46 44 4a 47 59
##  [64] 4b 57 4a 33 2f 64 6f 77 6e 6c 6f 61 64 2f 61 64 76 61 6e 63 65
##  [85] 64 74 65 63 68 6e 69 71 75 65 73 77 69 74 68 61 77 73 67 6c 75
## [106] 65 65 74 6c 6a 6f 62 73 5f 5f 31 5f 2e 70 64 66 50 08 0b 79 00
## [127] 00 00 00 00 00 01 01 00 00 00 00 00 00 00 03 00 00 00 00 00 00
## [148] 00 00 00 00 00 00 00 00 00 7a
```

There is a "string" version of the function, but it may return "nothing" if there are embedded NULLs or other breaking characters in the contents:

```{r 3, eval=FALSE}
get_xattr(sample_file, "com.apple.metadata:kMDItemWhereFroms")
## [1] "bplist00\xa2\001\002_\020khttps://files.slack.com/files-pri/T3V1ZDQHM-FDJGYKWJ3/download/advancedtechniqueswithawsglueetljobs__1_.pdfP\b\vy"
```

You are really better off doing this if you really want a raw string conversion:

```{r 4, eval=FALSE}
readBin(get_xattr_raw(sample_file, "com.apple.metadata:kMDItemWhereFroms"), "character")
## [1] "bplist00\xa2\001\002_\020khttps://files.slack.com/files-pri/T3V1ZDQHM-FDJGYKWJ3/download/advancedtechniqueswithawsglueetljobs__1_.pdfP\b\vy"
```

More often than not (on macOS) extended attributes are "binary property lists" (or "binary plist" for short). You can test to see if the returned raw vector is likely a binary plist:

```{r 5, eval=FALSE}
is_bplist(get_xattr_raw(sample_file, "com.apple.metadata:kMDItemWhereFroms"))
## [1] TRUE
```

If it is, you can get the data out of it. For now, this makes a system call to `plutil` on macOS and `plistutil` on other systems. You'll be given a hint on how to install `plistutil` if it's not found.

```{r 6, eval=FALSE}
read_bplist(get_xattr_raw(sample_file, "com.apple.metadata:kMDItemWhereFroms"))
## $plist
## $plist$array
## $plist$array$string
## $plist$array$string[[1]]
## [1] "https://files.slack.com/files-pri/T3V1ZDQHM-FDJGYKWJ3/download/advancedtechniqueswithawsglueetljobs__1_.pdf"
## 
## 
## $plist$array$string
## list()
## 
## 
## attr(,"version")
## [1] "1.0"
```

This is R, so you should really consider doing this instead of any of the above #rectanglesrule:

```{r 7, eval=FALSE}
get_xattr_df(sample_file)
## # A tibble: 2 x 3
##   name                                  size contents   
##   <chr>                                <dbl> <list>     
## 1 com.apple.metadata:kMDItemWhereFroms   157 <raw [157]>
## 2 com.apple.quarantine                    56 <raw [56]> 
```

you can live dangerously even with data frames, tho:

```{r 8, eval=FALSE}
get_xattr_df(sample_file) %>% 
  mutate(txt = map_chr(contents, readBin, "character")) # potentially "dangerous"
## # A tibble: 2 x 4
##   name                 size contents  txt                            
##   <chr>               <dbl> <list>    <chr>                          
## 1 com.apple.metadata…   157 <raw [15… "bplist00\xa2\x01\x02_\x10khtt…
## 2 com.apple.quaranti…    56 <raw [56… 0083;5bc9c8ca;Slack;44698902-A…
```

### Extended Example

We can process a whole directory of files to see what extended attributes various path targets have:

```{r 9, eval=FALSE}
fils <- list.files("~/Downloads", full.names = TRUE) 

xdf <- map_df(set_names(fils, fils), get_xattr_df, .id="path")

count(xdf, name, sort=TRUE) 
## # A tibble: 11 x 2
##    name                                                     n
##    <chr>                                                <int>
##  1 com.apple.quarantine                                    55
##  2 com.apple.metadata:kMDItemWhereFroms                    46
##  3 com.apple.lastuseddate#PS                               13
##  4 com.apple.metadata:_kMDItemUserTags                      3
##  5 com.apple.metadata:kMDItemDownloadedDate                 2
##  6 com.apple.diskimages.fsck                                1
##  7 com.apple.diskimages.recentcksum                         1
##  8 com.apple.FinderInfo                                     1
##  9 com.apple.metadata:com_apple_mail_dateReceived           1
## 10 com.apple.metadata:com_apple_mail_dateSent               1
## 11 com.apple.metadata:com_apple_mail_isRemoteAttachment     1
```

And we can work with `com.apple.metadata:kMDItemWhereFroms` binary plist data in bulk:

```{r 10, eval=FALSE}
filter(xdf, name == "com.apple.metadata:kMDItemWhereFroms") %>%
  filter(map_lgl(contents, is_bplist)) %>% 
  mutate(converted = map(contents, read_bplist)) %>% 
  select(size, converted) %>% 
  mutate(converted = map(converted, ~flatten_chr(.x$plist$array$string))) %>% 
  unnest() %>% 
  mutate(converted = urltools::domain(converted)) # you don't rly need to see the full URLs for this example
## # A tibble: 44 x 2
##     size converted                                              
##    <dbl> <chr>                                                  
##  1   189 dl.packetstormsecurity.net                             
##  2   129 files.slack.com                                        
##  3   120 files.slack.com                                        
##  4   117 files.slack.com                                        
##  5   592 irma.nps.gov                                           
##  6   157 files.slack.com                                        
##  7   624 github-production-release-asset-2e65be.s3.amazonaws.com
##  8   197 images.unsplash.com                                    
##  9   197 images.unsplash.com                                    
## 10   185 files.slack.com                                        
## # … with 34 more rows
```

### Full Suite

```{r full-suite}
# Create a temp file for the example
tf <- tempfile(fileext = ".csv")
write.csv(mtcars, tf)

# has attributes? (shld be FALSE)
has_xattrs(tf)
get_xattr(tf, "is.rud.setting")

# set an attribute
set_xattr(tf, "is.rud.setting.a", "first attribut")
get_xattr(tf, "is.rud.setting.a")
get_xattr_size(tf, "is.rud.setting.a")

# shld be TRUE
has_xattrs(tf)

set_xattr(tf, "is.rud.setting.b", "second attribute")
get_xattr(tf, "is.rud.setting.b")
get_xattr_size(tf, "is.rud.setting.b")

# overwrite an attribute
set_xattr(tf, "is.rud.setting.a", "first attribute")
get_xattr(tf, "is.rud.setting.a")
get_xattr_size(tf, "is.rud.setting.a")

# see all the attributes
list_xattrs(tf)

# data frame vs individual functions
get_xattr_df(tf)

# remove attribute
rm_xattr(tf, "is.rud.setting")
get_xattr(tf, "is.rud.setting")

# cleanup
unlink(tf)
```

## xattrs Metrics

```{r cloc, echo=FALSE}
cloc::cloc_pkg_md()
```

## Code of Conduct

Please note that this project is released with a Contributor Code of Conduct. By participating in this project you agree to abide by its terms.
